The limits of self commitment and private paternalism \thanks{}

### 2020-06-16 main.tex
copy repo

We contribute to several strands of literature. First, to the broad literature on commitment demand. A significant literature has shown that demand for commitment exists but is somewhat low in some contexts (\cite{Ashraf}, \cite{Gine}, \cite{Ted}), and high in others (\cite{Kremer},  \cite{Casaburi}, \cite{Alcohol}, \cite{AprajitP&P}). We add to this literature by showing that very small fees are enough to drastically reduce demand for commitment, even when commitment is beneficial for almost everybody in our sample, and that softer commitments in the form of non-pecuniary personal promises elicit higher demand.\footnote{Our fees are \hl{400\%} higher as a fraction of the loan than those of \cite{John}.} Second, two recent papers show that among those who demand commitment, the extent of commitment needed is underestimated (\cite{Ted}, \cite{John}) and that subjects can hurt themselves by self-imposing too week a commitment. We contribute in design ---by not allowing choice of the extent of commitment--- but also on how we infer the underestimation of the value of commitment by subjects. While these two papers infer too low commitment from subjects failing to reach the goal ex-post, we infer it by quantifying forgone personalized treatment effects ex-ante. Third, we focus on an important new context: pawnshop borrowing. Pawnshop borrowing is extremely prevalent in developing countries and very worrisome to regulators, yet severely understudied.

Fourth, we study \textit{both} causal effects and selection into frequent payment commitment loan contracts. \cite{Ghatak} state that ``high-frequency repayment, a feature in nearly all microfinance contracts, has been largely overlooked''.\footnote{\cite{LittleAtAtime} also note that ``it is especially surprising that there is so little analytical discussion of the incentive effects of the installment system''.} This paper fills part of this gap. \cite{Pande} study the effect of installment frequency on loan default and (surprisingly given their prevalence and the literature above) find null effects, while \cite{Murdoch} study selection into frequent payment microfinance loans, conjecturing that ``such repayment structure is potentially beneficial in fostering discipline among present-biased borrowers''. In contrast to \cite{Pande} we find very strong effects of payment frequency on decreased default.\footnote{Possibly because there are high levels of default in our context and not in theirs and because we have more statistical power from larger sample and individual randomization.} \cite{Murdoch}  show that a robust positive correlation exists between having present biased preferences in surveys and selecting into microfinance ---which they interpret as being due to the discipline of frequent payment--- acknowledging that the ``interpretation is suggestive... and needs to be confirmed by studies that establish causal links''. Our experimental test of this causal link rejects that people like and demand frequent payments (at least in our context). In contrast to \cite{Murdoch} we can isolate the pure effect of payment frequency on demand by randomly changing just this aspect of the contract. Fifth, we also contribute to the literature on the effect of promises by testing their effect on the field.

More importantly, our unique experimental design, paired with new machine learning methods to estimate heterogeneous treatment effects \cite{atheygrf}, allows us to estimate whether there is positive or negative selection into beneficial contracts, and whether pawnshop clients forgo money by not demanding commitment. In consonance with what \cite{Walters} finds in another context, we find that it is not the case that those that benefit most from the contract demand it, consistent with a significant fraction of clients displaying naivete about self control problems (see \cite{Rabin2018}). Our context is one in which private paternalism (\cite{Laibson2018}) could work since almost all pawnshop clients would benefit from frequent payment contract yet there it almost no demand for them. After our experiment ended, Lender P started to tie in installment payment components to some of their contracts themselves.

@@

The paper is organized as follows. Section \ref{context} provides context and describes the data sources. Section \ref{Experiment} describes the experiment and shows pre-treatment balance across arms. Section \ref{fee-commitment} estimates the effect of forcing clients into the fee-commitment contract, while Section \ref{fee-choice} studies the effects of choice. Section \ref{promises} assesses if promises work in our context as commitment devices. Finally, Section \ref{conclusion} concludes.

@@

Pawn borrowing is one of the oldest form of borrowing and one to which people with little access to the formal credit market turn to for ``emergency'' loans. They existed in antiquity at least since the Roman Empire, and there are records of it in China about 1,500 years ago (\cite{PawnShops}). In spite of this, there are extremely few studies in the economics literature about it. Why people go, how much it costs, what different contracts can tell us about consumer biases. We study such  context with Lender $P$, one of the largest pawn shops in Mexico, with dozens of branches spanning multiple States in Mexico.

Lender $P$ was interested in understanding why half of their clients lost their pawn. It was concerned with high default rates for two reasons. The first is that the pawned piece may have a sentimental value to the client above of its gold value, making it inefficient for the lender to keep the piece. We actually find that clients report a subjective value that is almost twice the value of its gold content. Second, it thought that keeping the piece lowered client satisfaction and made the client less likely to return and pawn again. They were interested in investigating whether a frequent payment contract would increase pawn recovery and costumer satisfaction, and whether there was demand for such a contract. Default is really  in this context  compared to many papers in the microfinance literature. \cite{Pande} study of the effect of payment frequency work in a context where the rate of default is close to 1\%. %\cite{LittleAtAtime} has a baseline default of \hl{XX\%}.

One potential explanation for high default is that clients are really just selling their gold piece. This is unlikely for the large majority of people not only because the reported subjective value of the pawn is larger than the loan size for 86\% of them, but also because among those that lose their pawn 74\% paid a positive amount towards its recovery. This can only be rationalized if they expected to recover their pawn.  Figure \ref{proxy_naive} focuses on those that ended up losing the pawn in our experiment and plots different measures of engagement with Lender $P$. Panel (c) shows total payments that went towards the loan. On average clients that lost their pawn paid 34\% of the value of their loan. Panel (d) shows that while 40\% made zero payments, the rest made one or several payments. In fact, panel (b) shows that close to 55\% of clients extended the loan for another cycle and made payments in the second cycle.

@@

Their business model is very simple. They take gold jewelry as collateral in exchange for a fraction of money given instantly in cash. The transaction takes less than 10 minutes and is conducted at the branch in person between the client and the appraiser (i.e. a teller). The appraiser weights the gold piece and runs tests on its purity. Based on these he assigns a gold value to the piece, stores it as collateral, and gives 70\% of the gold value of the piece instantly and in cash to the costumer. The costumer signs a 2-sheet contract with the main conditions of the contract at the branch and leaves with the cash.

Lender $P$ had only one type of contract. It stipulated that the interest rate was 7\% \textit{per month} accumulated daily on the outstanding amount of the loan. The loan had a 90 days term with 15 days grace period, and the client could make payments at anytime and no penalty for pre-payment. If the client paid the principal plus the accumulated interest before those 105 days, then she received back her pawn, otherwise the pawnbroker kept the piece and any payments done. Before the contract expired, the client had the right to renew it (using the same collateral) for another 3 months by going to the pawnshop paying the accumulated interest and signing a new contract with exactly the same terms. The pawnshop therefore makes money in three ways: by reselling the jewelry left as collateral on defaulted loans, by charging interest on non defaulted loans, and by keeping the payments made on defaulted loans.

This contract was standard in the industry, the clients that pawned understood its conditions (as we verified in interviews) and still decided to pawn. Bear in mind that these clients have little or no access to other types of loans, and also that they value its convenience: the process is fast and requires no documentation or credit-history. Potentially they like it also because they are overconfident on average about the probability of recovering their pawn, as discussed below, even though 90\% of them have pawned before.

\section{Data and summary stats}

@@

\subsection{Administrative Data}
The administrative data contains a unique identifier of the client, an identifier of the piece she is pawning, and the transactions relating to that piece. In particular, the value of the piece as assessed by the appraiser, the amount of money lent (70\% of the value), and the date when it was pawned. As described below, for our experiment we also recorded the type of contract for that pawn. After the pawn happens, we follow each transaction related to that piece in the administrative data: when were payments made and for which amounts, whether there was default (i.e. the client lost her pawn), and whether any late-payment fees (active for only some experimental arms) were imposed. We have this information for all the pawns that occurred in our 6 branches between August 2nd 2012 and August 13th 2013, this includes all the pawns under our experiment but also those that happened 1 month before it started and up to 8 months after it ended. Figure \ref{exp_description}(a) shows the timing of the experiment and our admin data coverage. The black arrow represent the dates for which we have administrative data (before and after the experiment) while the light gray represents the duration of the experiment. The experiment comprises 13,444 pawns, and our administrative data cover a total of 26,179 pawns.

@@

We had a team of enumerators in each branch collect surveys and clearly explain the contract terms to walk-in clients. The enumerators were inside the branch and asked clients to complete a short 5 minutes survey before going to the teller window to appraise their piece. The survey was necessarily short to avoid discouraging the potential clients from pawning, but at the same time it aimed to measure the following: demographics, proxies for income/wealth, education, self-control problems/present-biased preferences, experience pawning, if the family or friends commonly asked for money, how time consuming and costly was to come to the branch, the subjective probability of recovering their piece that they intended to pawn, the subjective value of their piece in money terms (for how much money would they sell it for), among others. We were able to survey 7,704 clients (some refused). %\footnote{We also surveyed clients before and after the experiment, so our survey covers a larger sample that we are not using in this paper.}. 
The appendix \ref{baseline_survey} transcribes the questionnaire in English and presents summary stats. Survey response rate is 57\%, the Appendix Table \ref{balance_response} shows that response is not related to admin data observables.

@@

Table \ref{SS} presents some summary statistics for the whole sample and by treatment arm. Because treatment arms are described below, here we concentrate on the overall average. Panel A describes variables from the administrative data. It shows that the average loan size is \$1945 pesos (\$105 usd). The average number of pawns per day per branch is 33. Although not shown in the table, only 40\% of clients recover they pawn within 230 days. Even those that recover their pawn tend to pay it back at the last moment, only 17\% pay before the 90th day.

Panel B reports stats from our survey data. We learn that 73\% of clients are women, with an average age of 44 years; 66\% of them have completed high school or more, and 90\% have pawned before, so that our sample has mostly experienced borrowers. Finally the subjective probability of recovery is close to 93\% on average, which contrast with the 44\% recovery that happens in reality for the pawns in the control group. Just by this measure borrowers seem to be highly overconfident on average. It turns out that the average subjective value they report for the pieces is 3250 pesos, which is much larger than the average appraised gold value (1950 pesos). This makes it inefficient for them to lose it.

Since the main treatment involves making them come to the branch to pay each month, the transaction costs of coming to the branch are important, and we ask two questions around them. The average time they take to come to the branch is 22 minutes, and the amount of money they spend in transport to do that is 11 pesos (\$0.60 usd).  
The population we are working with is economically vulnerable in the sense that 31\% of them could not pay either water, electricity \& gas or rent in the past 6 months. They often receive negative income shocks: 87\% said they are pawning because of an emergency, and only 13\% stated it was to use in a `non-urgent expense'.
%When asked why they are pawning this piece \hlgr{5\%} responded \hlgr{`lost a family member'}, \hlgr{`a medical emergency'} (\hlgr{11\%}), \hlgr{`an urgent expense'} (\hlgr{71\%}),

@@

Our main dependent variables are loan default and financial cost incurred by the borrower. Default is defined as losing the pawn. We observe directly when the piece gets appropriated by Lender $P$ and goes on sale.

We define financial cost as the discounted sum of payments toward the loan plus the value of the loan, and substract to it the appraised or subjective discounted value of the pawn when it is recovered: $Financial \; Cost_i = \sum_t \frac{P_{it}}{(1+r)^t} + Pawn \: Value_i - \frac{\text{I}(Recovered \: Pawn_i) \times Pawn \: Value_i}{(1+r)^T}$, with $t$ indexing days, $T$ is the day when the pawn is recovered, I($\cdot$) an indicator function for recovering the pawn, $r$ is a daily interest rate equivalent to a 7\% monthly interest rate. Figure \ref{fc_hist}(a) plots a histogram of this cost in pesos conditional on recovering and conditional on not recovering the pawn. This cost is calculated for a time period of 230 days after loan origination, as 50\% of clients actually renew the loan one time and 28\% a second time. Of those who renew once 34\% lose the pawn, and of those that renew twice 15\% do. Figure \ref{fc_hist}(a) suggest that financial cost is significant, even for those that do recover their pawn.  Panel (b) normalizes the cost by loan size. It shows that as a percentage of their pawn, most of the cost comes from clients that do not recover their pawn. For them the cost is between 150\% and 250\% of the loan value. %\hl{The difference between panel (a) and (b) arises since clients who lose their pawn tend to be those that pledge cheaper pieces.} 
We calculate an APR of \hl{177\% on average for the control group.}\footnote{The Annual Percentage Rate (for pawn $j$) is calculated as the internal rate of return $i$ such that $\sum_t \frac{P_{jt}}{(1+i)^t} + \frac{\text{I}(Default \: Pawn_j) \times Pawn \: Value_j}{(1+i)^T} - Pawn \: Value_j = 0$, where $T$ is the date the pawn was lost.}

@@

\noindent \textit{Selection of branches.} Starting September 6th 2012, we implemented our experiment in 6 branches of Lender $P$. The branches were selected by Lender $P$ to be dispersed across Mexico city and have varying sizes. In four of them the experiment ran for 107 days, and in 2 of them we ran it for a shorter time to economize on data collection costs since we realized we would not be constrained by sample size. %Thus we eliminated \hl{the smallest} ones. 
Branches are more than 5 km apart from each other, and there is little substitution among them (only 1\% of consumers appear in more than one of our branches).

\vspace{.2in}
\noindent \textit{The monthly payment contract.} For the purpose of the experiment we designed a new contract that is identical to the status quo contract in all respects except one. It has the same interest rate (7\% \textit{per month} which accumulates daily on outstanding debt, it has the same loan size/collateral ratio (70\%), it has the same loan term (90 days), and the gold pawn gets appraised in the same way by the same appraisers. The only difference is that it has a requirement that the client has to pay monthly: for the 3 month duration at least 1/3 of the loan value (plus accumulated interest) has to be paid before day 30th, 60th and 90th after loan disbursement. Failure to pay these amount triggers a penalty fee of 2\% of this minimum amount due (in some of the treatment arms), which is added to the amount that the client needs to pay to recover their pawn. This penalty acts as a commitment device to pay now vs paying later. The level of the penalty was determined by Lender $P$ after talking to clients and based on Lender $P$'s experience. They deemed 2\% something that was big enough to provide commitment but small enough that if clients had an emergency and could not pay, the fee would not cause the pawn to be unrecoverable.\footnote{As explained below we also experimented with a contract where there was no pecuniary fee for paying late but where we clients made a personal promise to pay monthly. In contrast with \cite{John} we don't let the clients choose the size of the fee. As that paper shows doing this results in wrong choices by naive present biased consumers.}

\vspace{.2in}
\noindent \textit{Randomization.} Randomization was done at the branch-day level.  Each day the system chose which types of contracts were on offer that day in the branch, and therefore which contract the client signed. Branch personnel did not know which treatment would happen which day. They were told that there were 5 different ``types of contract-days'' and that the system chose randomly which applied any given date, and that it could happen for instance that two consecutive dates had the same contract. They were also told that this way of operating was in effect in various Lender $P$'s branches (they did not know which ones), and that it could be in place for several months. Randomizing at the day level limits the problem of contamination arising from clients waiting for appraisal observing that other clients get different contracts than theirs. It also limits potential  manipulation by appraisers, who in the presence of individual level randomization could potentially pick their preferred costumers from the line or tell them to wait until their desired contract shows up on the screen. Intra-branch correlation on the probability of default (ICC) is small, at 0.08, so we don't lose much power vis-a-vis individual level randomization.

\vspace{.2in}
\noindent \textit{Treatment Arms.} We have 5 different experimental arms, which differ in which contracts were offered on a given branch-day.

\begin{enumerate}
    \item \textit{Status-quo} arm: consisted of branch-days offering the status quo contract described in Section \ref{context}, and only this contract (as was standard). 
    \item \textit{Fee-Forcing} arm: consisted of branch-days offering only the monthly payment contract with a fee for late payment. 
    \item \textit{Fee-Choice} arm: consisted of branch-days where the client was given \textit{a choice} between the monthly payment with fee contract, and the status quo contract.
    \item \textit{Promise-Forcing} arm: consisted of branch-days which offered the monthly payment contract only, but where there was no fee penalty. Instead, the client was made to sign a paper which said ``I promise to pay every month the corresponding sum of \_\_\_\_\_\_, on the dates \_\_\_\_\_\_, \_\_\_\_\_\_, and \_\_\_\_\_\_. This is \underline{not} a legal document and cannot be used in courts. It is just a \textit{personal promise}. If I do not comply I would have broken my word.''. After signing, the promise was read to her by the appraiser.
    \item \textit{Promise-Choice} arm: consisted on branch-days where the client was given \textit{a choice} between the Promise-Forcing and the Status quo contract.\footnote{We did not allocate equal number of days across arms, since we were interested in having more power in some of them. The days allocated to each -- in order-- were 84, 80, 93, 68, 82.}
\end{enumerate}

By comparing the \textit{Fee-Forcing} group vs \textit{Status-quo} group we estimate the causal effect of the commitment contract over the contract where the client can paying whenever they want in a 105 day period. On the other hand, having arms where clients could choose among these contracts allows us to study their demand for commitment. Note that a neoclassical consumer (with no intra-household concerns) would not voluntarily restrict her options of payment profiles (as in monthly payment contract) and be subject to fees since she can always replicate the monthly payment profile herself under the status quo contract if she so desired. The status quo contract gives her extra flexibility to respond to income shocks for instance. However, if clients have present biased preferences \textit{and} are sophisticated about them, they may choose the frequent payment contract as it has commitment value. This value would be traded off against the value of flexibility. This design also allows us to test if promises work in a natural field environment which to our knowledge has not been done. Promises could have larger or smaller commitment value that a pecuniary fee, something that we test by comparing behavior in the \textit{Promise-Forcing}  vs \textit{Status-quo} arms.

@@

We are fairly confident that the overwhelming majority of clients understood the contracts and made informed choices. As shows below, different contract terms did generate differential demand and differential behavior. Moreover, we can systematically predict demand based on consumer characteristics, beliefs and proxies for present biased preferences.

\section{The effect of forcing into fee-commitment} \label{fee-commitment}

@@

\subsubsection{Behavioral economics explanations}

@@

>>>>

We contribute to several strands of literature. First, to the broad literature on commitment demand. A significant literature has shown that demand for commitment exists but is somewhat low in some contexts (\cite{Ashraf}, \cite{Gine}, \cite{Ted}), and high in others (\cite{Kremer},  \cite{Casaburi}, \cite{Alcohol}, \cite{AprajitP&P}). We add to this literature by showing that very small fees are enough to drastically reduce demand for commitment, even when commitment is beneficial for almost everybody in our sample, and that softer commitments in the form of non-pecuniary personal promises elicit higher demand.\footnote{Our fees are 16 times smaller than than those of \cite{John} as a fraction of the loan/goal, and five times smaller converting peso values using exchange rates.} Second, two recent papers show that among those who demand commitment, the extent of commitment needed is underestimated (\cite{Ted}, \cite{John}) and that subjects can hurt themselves by self-imposing too week a commitment. We contribute in design ---by not allowing choice of the extent of commitment--- but also on how we infer the underestimation of the value of commitment by subjects. While these two papers infer too low commitment from subjects failing to reach the goal ex-post, we infer it by quantifying forgone personalized treatment effects ex-ante. Third, we focus on an important new context: pawnshop borrowing. Pawnshop borrowing is extremely prevalent in developing countries and very worrisome to regulators, yet severely understudied.

Fourth, we study \textit{both} causal effects and selection into frequent payment commitment loan contracts. \cite{Ghatak} state that ``high-frequency repayment, a feature in nearly all microfinance contracts, has been largely overlooked''.\footnote{\cite{LittleAtAtime} also note that ``it is especially surprising that there is so little analytical discussion of the incentive effects of the installment system''.} This paper fills part of this gap. \cite{Pande} study the effect of installment payment frequency on loan default and (surprisingly given their prevalence and the literature above) find null effects. \cite{Murdoch} study selection into frequent payment microfinance loans, conjecturing that ``such repayment structure is potentially beneficial in fostering discipline among present-biased borrowers'' and may explain why borrowers choose it. In contrast to \cite{Pande} we find strong effects of payment frequency on decreased default.\footnote{Possibly because there are high levels of default in our context and not in theirs and because we have more statistical power from a larger sample and individual level randomization.} \cite{Murdoch}  show that a robust positive correlation exists between displaying present biased preferences in surveys and selecting into microfinance ---which they interpret as being due to high demand for the discipline of frequent payments--- acknowledging that the ``interpretation is suggestive... and needs to be confirmed by studies that establish causal links''. Our experimental test of this causal link rejects that most people like and demand frequent payments (at least in our context). In contrast to \cite{Murdoch} we can isolate the pure effect of payment frequency on demand by randomly changing just this aspect of the contract. Fifth, we also contribute to the literature on the effect of promises by testing their effect on the field.

More importantly, our unique experimental design which combines forcing and choice arms, paired with new machine learning methods to estimate heterogeneous treatment effects (\cite{atheygrf}), allows us to estimate whether there is positive or negative selection into beneficial contracts, and whether pawnshop clients forgo money by not demanding commitment. In consonance with what \cite{Walters} finds in another context, we find that it is not the case that those that benefit most from the contract demand it more, consistent with a significant fraction of clients displaying naivete about self control problems (see \cite{Rabin2018}). Our context is one in which private paternalism (\cite{Laibson2018}) could work since almost all pawnshop clients would benefit from frequent payment contract yet there it almost no demand for them. After our experiment ended, Lender P started to tie in installment payment components to some of their contracts themselves.

@@

The paper is organized as follows. Section \ref{context} provides context and describes the data sources. Section \ref{Data} presents summary statistics and defines our main outcome variables. Section \ref{Experiment} describes the experiment and shows pre-treatment balance across arms. Section \ref{fee-commitment} estimates the effect of forcing clients into the fee-commitment contract, while Section \ref{fee-choice} studies the effects of choice. Section \ref{model} collects all findings and presents a model to rationalize them qualitatively. Section \ref{promises} assesses if promises work in our context as commitment devices. Finally, Section \ref{conclusion} concludes.

@@

Pawn borrowing is one of the oldest forms of borrowing and one to which people with little access to the formal credit market turn to for ``emergency'' loans. They existed in antiquity at least since the Roman Empire, and there are records of it in China about 1,500 years ago (\cite{PawnShops}). As explained above pawn loans involve individuals leaving valuable and liquid assets (typically jewelry) as collateral in exchange for an immediate cash loan. We study such  context by partnering with Lender $P$, one of the largest pawn shops in Mexico, with more than one hundred branches spanning multiple States in Mexico.

Lender $P$ was interested in understanding why half of their clients lost their pawn. Default is really in this context  compared to many papers in the microfinance literature\footnote{\cite{Pande} study of the effect of payment frequency in a context where the rate of default is close to 1\%.} It was concerned with high default rates for two reasons. The first is that the pawned piece may have a sentimental value to the client above of its gold value, making it inefficient for the lender to keep the piece. We actually find that clients report a subjective value that is almost twice the value of its gold content. Second, keeping the piece may lower client satisfaction and make the client less likely to return and pawn again. They were interested in investigating whether a frequent payment contract would increase pawn recovery and costumer satisfaction, and whether there was demand for such a contract.

One potential explanation for high default is that clients are really just selling their gold piece. This is unlikely for severalreasons: (a)  the reported subjective value of the pawn is larger than the loan size for \hl{86\%} of them, (b) among those that lose their pawn 74\% paid a positive amount towards its recovery --- this can only be rationalized if they expected to recover their pawn, (c) it is typical to find gold purchasing stores close to where the pawnshops are located (see Figure \ref{GoldBuyers}) and they give more than what Lender P gives for the pawn, (d) there is a lot of engagement with the lender after pawning even for those who lost their pawn. On average clients that lost their pawn paid 34\% of the value of their loan, and close to 55\% of clients extended the loan for another cycle and made payments in the second cycle (see Figure \ref{proxy_naive} in Appendix).

@@

Lender P's business model is very simple. They take gold jewelry as collateral in exchange for a fraction of money given instantly in cash. The transaction takes less than 10 minutes and is conducted at the branch in person between the client and the appraiser (i.e. a teller, see Figure \ref{PawnshopPicture}). The appraiser weights the gold piece and runs tests on its purity. Based on these he assigns a gold value to the piece, stores it as collateral, and gives 70\% of the gold value of the piece instantly and in cash to the costumer. The costumer signs a 2-sheet contract with the main conditions of the contract at the branch and leaves with the cash.

Lender $P$ had only one type of contract. It stipulated that the interest rate was 7\% \textit{per month} accumulated daily on the outstanding amount of the loan. The loan had a 90 days term with 15 days grace period, and the client could make payments at anytime with no penalty for pre-payment. If the client paid the principal plus the accumulated interest before those 105 days, then she received back her pawn, otherwise the pawnbroker kept the piece and any payments done. Before the contract expired, the client had the right to renew it (using the same collateral) for another 3 months by going to the pawnshop paying the accumulated interest and signing a new contract with exactly the same terms as the original one. %The pawnshop therefore makes money in three ways: by reselling the jewelry left as collateral on defaulted loans, by charging interest on non defaulted loans, and by keeping the payments made on defaulted loans.

This contract was standard in the industry, the clients that pawned understood its conditions (as we verified in interviews) and still decided to pawn.\footnote{90\% of clients report in our survey they have pawned before.} Bear in mind that these clients have little or no access to other types of loans, and also that they value its convenience: the process is fast and requires no documentation or credit-history. Potentially they like it also because they are overconfident on average about the probability of recovering their pawn, as discussed below.

\section{Data and summary stats} \label{Data}

@@

\subsection{Administrative Data}
The administrative data contains a unique identifier of the client, an identifier of the piece she is pawning, and the transactions relating to that piece. In particular, the value of the piece as assessed by the appraiser, the amount of money lent (70\% of the value), and the date when it was pawned. As described below, for our experiment the administrative data recorded the type of contract for that pawn. After the pawn happens, we follow each transaction related to that piece in the administrative data as well as when were payments made and for which amounts, whether there was default (i.e. the client lost her pawn), and whether any late-payment fees (active for the relevant experimental arms) were imposed. We have this information for all the pawns that occurred in our 6 branches between August 2nd 2012 and August 13th 2013, this includes all the pawns under our experiment but also those that happened 1 month before it started and up to 8 months after it ended. Figure \ref{exp_description}(a) shows the timing of the experiment and our admin data coverage. The black arrow represent the dates for which we have administrative data (before and after the experiment) while the light gray represents the duration of the experiment. The experiment comprises 13,444 pawns, and our administrative data cover a total of 26,179 pawns.

@@

We had a team of ITAM student enumerators in each branch collect surveys and clearly explain the contract terms to walk-in clients. The enumerators were inside the branch and asked clients to complete a short 5 minutes survey before going to the teller window to appraise their piece. The survey was necessarily short to avoid discouraging the potential clients from pawning, but at the same time it aimed to measure the following: demographics, proxies for income/wealth, education, self-control problems/present-biased preferences, experience pawning, if the family or friends commonly asked for money, how time consuming and costly was to come to the branch, the subjective probability of recovering their piece that they intended to pawn, the subjective value of their piece in money terms (for how much money would they sell it for), among others. We were able to survey 7,704 clients (some refused). %\footnote{We also surveyed clients before and after the experiment, so our survey covers a larger sample that we are not using in this paper.}. 
The appendix \ref{baseline_survey} transcribes the questionnaire in English and presents summary stats. Survey response rate is 57\%, the Appendix Table \ref{balance_response} shows that response is not related to admin data observables.

@@

Table \ref{SS} presents some summary statistics for the whole sample and by treatment arm. Because treatment arms are described below, here we concentrate on the overall average. Panel A describes variables from the administrative data. It shows that the average loan size is \$1945 pesos (\$105 usd). The average number of pawns per day per branch is 33. Although not shown in the table, only 40\% of clients recover they pawn within 230 days. Even those that recover their pawn tend to pay it back at the last moment, with only 17\% paying before the 90th day.

Panel B reports stats from our survey data. We learn that 73\% of clients are women, with an average age of 44 years; 66\% of them have completed high school or more, and 90\% have pawned before, so that our sample has mostly experienced borrowers. Finally the subjective probability of recovery is close to 93\% on average, which contrast with the 44\% recovery that happens in reality for the pawns in the control group. Just by this measure borrowers seem to be highly overconfident on average (more on this below). It turns out that the average subjective value they report for the pieces is 3250 pesos, which is much larger than the average appraised gold value of 1950 pesos. This makes it inefficient for them to lose it.

Since the main treatment involves making clients come to the branch to pay each month, the transaction costs of coming to the branch are important, and we ask two questions about them. The average time they take to come to the branch is 22 minutes, and the amount of money they spend in transport to do that is 11 pesos (\$0.60 usd).  
The population we are working with is economically vulnerable in the sense that 31\% of them could not pay either water, electricity \& gas or rent in the past 6 months. They often receive negative income shocks: 87\% said they are pawning because of an emergency, and only 13\% stated it was to use in a `non-urgent expense'.
%When asked why they are pawning this piece \hlgr{5\%} responded \hlgr{`lost a family member'}, \hlgr{`a medical emergency'} (\hlgr{11\%}), \hlgr{`an urgent expense'} (\hlgr{71\%}),

@@

Our main dependent variables are loan default and financial cost incurred by the borrower. Default is defined as losing the pawn. We observe directly when the piece gets appropriated by Lender $P$.

We define financial cost as the discounted sum of payments toward the loan plus the value of the pawn, and subtract the subjective discounted value of the pawn when it is recovered: $Financial \; Cost_i = \sum_t \frac{P_{it}}{(1+r)^t} + Pawn \: Value_i - \frac{\text{I}(Recovered \: Pawn_i) \times Pawn \: Value_i}{(1+r)^T}$, with $t$ indexing days, $T$ is the day when the pawn is recovered, I($\cdot$) an indicator function for recovering the pawn, $r$ is a daily interest rate equivalent to a 7\% monthly interest rate.\footnote{We experimented with different interest rates with similar results. See section \ref{neoclasical}.} Figure \ref{fc_hist}(a) plots a histogram of this cost in pesos for loans where the pawn was recovered or lost. This cost is calculated for a time period of 230 days after loan origination, as 50\% of clients actually renew the loan one time. Of those who renew 34\% lose the pawn. Figure \ref{fc_hist}(a) suggest that financial cost is significant, even for those that do recover their pawn.  Panel (b) normalizes the cost by loan size. It shows that as a percentage of their pawn, most of the cost comes from clients that do not recover their pawn. For them the cost is between 140\% and 250\% of the loan value. %\hl{The difference between panel (a) and (b) arises since clients who lose their pawn tend to be those that pledge cheaper pieces.} 
We calculate an APR of \hl{177\% on average for the control group.}\footnote{The Annual Percentage Rate (for pawn $j$) is calculated as the internal rate of return $i$ such that $\sum_t \frac{P_{jt}}{(1+i)^t} + \frac{\text{I}(Default \: Pawn_j) \times Pawn \: Value_j}{(1+i)^T} - Pawn \: Value_j = 0$, where $T$ is the date the pawn was lost.}

@@

\noindent \textbd{Selection of branches.} Starting September 6th 2012, we implemented our experiment in 6 branches of Lender $P$. The branches were selected by Lender $P$ to be dispersed across Mexico city and have varying sizes. In four of them the experiment ran for 107 days, and in 2 of them we ran it for a shorter time to economize on data collection costs since we realized we would not be constrained by sample size. %Thus we eliminated \hl{the smallest} ones. 
Branches are more than 5 km apart from each other, and there is little substitution among them (only 1\% of consumers appear in more than one of our branches).

\vspace{.2in}
\noindent \textbf{The monthly installment contract.} For the purpose of the experiment we designed a new contract that is identical to the status quo contract in all respects except one. It has the same interest rate (7\% \textit{per month} which accumulates daily on outstanding debt, it has the same loan size/collateral ratio (70\%), it has the same loan term (90 days, and a grace period of 15 days), and the gold pawn gets appraised in the same way by the same appraisers. The only difference is that it has a requirement that the client has to pay monthly: for the 3 month duration at least 1/3 of the loan value (plus accumulated interest) has to be paid before day 30th, 60th and 90th after loan disbursement. Failure to pay this amount triggers a penalty fee of 2\% of this minimum amount due, which is added to the amount that the client needs to pay to recover their pawn. This penalty acts as a commitment incentive to pay now vs paying later. The level of the penalty was determined by Lender $P$ after talking to clients and based on Lender $P$'s experience. They deemed 2\% something that was big enough to provide commitment but small enough that if clients had an emergency and could not pay, the fee would not cause the pawn to be unrecoverable.\footnote{In contrast with \cite{John} we don't let the clients choose the size of the fee. As that paper shows doing this results in wrong choices by naive present biased consumers. Our fee is much smaller than the median in her experiment as explained above. We also experimented with a contract where there was no pecuniary fee for paying late but where we clients made a personal promise to pay monthly.}

\vspace{.2in}
\noindent \textbf{Randomization.} Randomization was done at the branch-day level.  Each day the system chose which types of contracts were on offer that day in the branch, and therefore which contract the client signed. Branch personnel did not know which treatment would happen which day. They were told that there were 5 different ``types of contract-days'' and that the system chose randomly which applied any given date, and that it could happen for instance that two consecutive dates had the same contract. They were also told that this way of operating was in place onin various Lender $P$'s branches (they did not know which ones), and that it could be in place for several months. Randomizing at the day level limits the problem of contamination arising from clients realizing that other clients get different contracts than theirs. It also limits potential  manipulation by appraisers, who in the presence of individual level randomization could potentially pick their preferred costumers from the line or tell them to wait until their desired contract shows up on the screen. Intra-branch correlation on the probability of default (ICC) is small, at 0.08, so we don't lose much power vis-a-vis individual level randomization.

\vspace{.2in}
\noindent \textbf{Treatment Arms.} We have 5 different experimental arms, which differ in which contracts were offered on a given branch-day.

\begin{enumerate}
    \item \textit{Status-quo forcing} arm: consisted of branch-days offering the status quo contract described in Section \ref{context}, and only this contract (as was standard). 
    \item \textit{Fee-Forcing} arm: consisted of branch-days offering only the monthly installment contract with a fee for late payment as described above. 
    \item \textit{Fee-Choice} arm: consisted of branch-days where the client was given \textit{a choice} between the monthly installment with fee contract, and the status quo contract.
    \item \textit{Promise-Forcing} arm: consisted of branch-days which offered the monthly payment contract only, but where there was no fee for paying late. Instead, the client was made to sign a paper which said ``I promise to pay every month the corresponding sum of \_\_\_\_\_\_, on the dates \_\_\_\_\_\_, \_\_\_\_\_\_, and \_\_\_\_\_\_. This is \underline{not} a legal document and cannot be used in courts. It is just a \textit{personal promise}. If I do not comply I would have broken my word.''. After signing, the promise was read to the client by the appraiser.
    \item \textit{Promise-Choice} arm: consisted on branch-days where the client was given \textit{a choice} between the promise-forcing and the status quo contract.\footnote{We did not allocate equal number of days across arms, since we were interested in having more power in some of them. The branch days allocated to each -- in order-- were 84, 80, 93, 68, 82.}
\end{enumerate}

By comparing the \textit{fee-forcing} arm vs the \textit{status-quo} arm we estimate the causal effect of mandatory installment frequency which acts as a incentive to pay earlier vis-a-vis the pay-when-you-want contract. On the other hand, having arms where clients can choose among these contracts allows us to study client's demand for commitment to pay earlier. Note that a neoclassical consumer (with no intra-household concerns) would not voluntarily restrict her options of payment profiles (as in monthly payment contract) by being subject to fees, since she can always replicate the monthly payment profile herself under the status quo contract if she so desired. The status quo contract gives her extra flexibility to respond to negative income shocks. However, if clients have present biased preferences \textit{and} are sophisticated about them, they may choose the frequent payment contract as it has commitment value. This value would be traded off against the value of flexibility. This design also allows us to test if promises work in a natural field environment which to our knowledge has not been done. Promises could have larger or smaller commitment value that a pecuniary fee, something that we test by comparing behavior in the \textit{promise-forcing}  vs \textit{status-quo} arms.

@@

We are fairly confident that the overwhelming majority of clients understood the contracts and made informed choices. As shown below, different contract terms did generate differential demand and differential behavior. Moreover, we can systematically predict demand based on consumer characteristics, beliefs and proxies for present biased preferences.

\section{The effect of forcing clients into fee-commitment} \label{fee-commitment}

@@

\subsubsection{Behavioral economics explanations} \label{behavioral}

@@

\subsection{Pictures}
